---
title: "hw1_833001533_ara"
author: "Anjana Raj A"
date: "2024-02-03"
output:
  pdf_document: default
---

```{r installing and running packages}
library("fGarch")
library("tigerstats")
library("ggplot2")
library("TTR")
library("gridExtra")
library("quantmod")
```

```{r Comparing Distirbutions}
x=seq(-50,50)
y1 =dsstd(x, mean=1, sd=1, xi=1)
y2 =dsstd(x, mean=1, sd=2, xi=1)
y3 =dsstd(x, mean=1, sd=1, xi=3)
y4 =dsstd(x, mean=4, sd=1, xi=1)
df1= data.frame( x=x, y1=y1, y2=y2, y3=y3, y4=y4, each=101)
p1=ggplot(df1, aes(x, y1)) + geom_line()+labs(title="Benchmark: with same mean, st. dev and mu")+ theme_minimal() +
  theme(plot.title = element_text(size = 8))
p2=ggplot(df1, aes(x, y2)) + geom_line()+labs(title="Changing st.deviation but keeping mean and mu constant")+theme_minimal() +
  theme(plot.title = element_text(size = 8))
p3=ggplot(df1, aes(x, y3)) + geom_line()+labs(title="Changing mu but keeping mean and st.deviation constant")+theme_minimal() +
  theme(plot.title = element_text(size = 8))
p4=ggplot(df1, aes(x, y4)) + geom_line()+labs(title="Changing mean but keeping mu and st.deviation constant")+theme_minimal() +
  theme(plot.title = element_text(size = 8))

tdist=grid.arrange(p1, p2, p3, p4, nrow = 2, ncol=2)
```
Increasing the sd flattens the distribution, decreasing the degrees of freedom xi makes it more peaked, and changing the mean shifts the distribution horizontally. The t-distribution shape is very sensitive to the xi parameter controlling the degrees of freedom.

```{r hedge fund}

#### Expected profit with loss $50k
niter = 2e5
Days = 100

profit = rep(0,niter)

for(j in c(1:niter)){
  daily = rnorm(Days, mean =  0.05/253, sd = 0.23/sqrt(253))
  closes = log(1e6) + cumsum(daily)

  minDay = which(closes < log(950000))[1]
  if(is.na(minDay)==TRUE) minDay = Days

  maxDay = which(closes > log(1100000))[1]
  if(is.na(maxDay) == TRUE) maxDay = Days
  
  saleDay = min(minDay,maxDay)
  profit[j] = exp(closes[saleDay]) - 1e6

}

hist(profit, 50)
exp_profit_50k=mean(profit)


#### Expected profit with loss $40k
niter = 2e5
Days = 100

profit = rep(0,niter)

for(j in c(1:niter)){
  daily = rnorm(Days, mean =  0.05/253, sd = 0.23/sqrt(253))
  closes = log(1e6) + cumsum(daily)

  minDay = which(closes < log(960000))[1]
  if(is.na(minDay)==TRUE) minDay = Days

  maxDay = which(closes > log(1100000))[1]
  if(is.na(maxDay) == TRUE) maxDay = Days
  
  saleDay = min(minDay,maxDay)
  profit[j] = exp(closes[saleDay]) - 1e6

}

hist(profit, 50)
exp_profit_40k=mean(profit)

#### Expected profit with loss $30k
niter = 2e5
Days = 100

profit = rep(0,niter)

for(j in c(1:niter)){
  daily = rnorm(Days, mean =  0.05/253, sd = 0.23/sqrt(253))
  closes = log(1e6) + cumsum(daily)

  minDay = which(closes < log(970000))[1]
  if(is.na(minDay)==TRUE) minDay = Days

  maxDay = which(closes > log(1100000))[1]
  if(is.na(maxDay) == TRUE) maxDay = Days
  
  saleDay = min(minDay,maxDay)
  profit[j] = exp(closes[saleDay]) - 1e6

}

hist(profit, 50)
exp_profit_30k=mean(profit)

##3 Changing Threshold from 100k

niter = 2e5
Days = 100

profit = rep(0,niter)

for(j in c(1:niter)){
  daily = rnorm(Days, mean =  0.05/253, sd = 0.23/sqrt(253))
  closes = log(12000000) + cumsum(daily)

  minDay = which(closes < log(970000))[1]
  if(is.na(minDay)==TRUE) minDay = Days

  maxDay = which(closes > log(1100000))[1]
  if(is.na(maxDay) == TRUE) maxDay = Days
  
  saleDay = min(minDay,maxDay)
  profit[j] = exp(closes[saleDay]) - 12000000

}

hist(profit, 50)
exp_profit_threshold=mean(profit)

#Change the number of trading days

niter = 2e5
Days = 40

profit = rep(0,niter)

for(j in c(1:niter)){
  daily = rnorm(Days, mean =  0.05/253, sd = 0.23/sqrt(253))
  closes = log(1e6) + cumsum(daily)

  minDay = which(closes < log(950000))[1]
  if(is.na(minDay)==TRUE) minDay = Days

  maxDay = which(closes > log(1100000))[1]
  if(is.na(maxDay) == TRUE) maxDay = Days
  
  saleDay = min(minDay,maxDay)
  profit[j] = exp(closes[saleDay]) - 1e6

}

hist(profit, 50)
trading_40=mean(profit)


expected_profits=as.table(c(loss50k=exp_profit_50k, loss40k=exp_profit_40k, loss30k=exp_profit_30k, thresholdChange=exp_profit_threshold, dayChange=trading_40))
expected_profits
```
Decreasing the lower loss threshold from 50K to 40K or 30K decreases expected profits.
Increasing profit threshold from 10k to 12k increases expected profits.
By reducing the trading days from 100 to 40, the expected loss declines.


```{r problem 4}
 niter = 10000
value = matrix(niter)
 set.seed(2015)
for (i in 1:niter)
{
Return = rnorm(20, mean =  0.002, sd = 0.03)
prices = 93.45 * exp(cumsum(Return))
ind = (min(prices) < 85)
value[i] = 100*ind - (3 - ind)
 }
mu = mean(value)
mu
```

```{r problem 5 and 6}
set.seed(123)  # Setting seed for reproducibility

# Parameters
mean_return =0.0002
sd_return = 0.01
initial_price = 93.45
strike_price_1 =85
strike_price_2 = 84.5
additional_payment = 200
iterations = 10000

# Simulate stock price movements
simulated_returns = rnorm(iterations, mean_return, sd_return)
simulated_prices = initial_price * exp(cumsum(simulated_returns))

# Calculate profit/loss
profit_loss =ifelse(simulated_prices < strike_price_1, 100, 
                      ifelse(simulated_prices < strike_price_2, 300, -3))

# Calculate mean profit/loss
mean_profit_loss = mean(profit_loss)

# Print the result
cat("Estimated mean profit/loss:", mean_profit_loss, "\n")


```



